---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading Required Libraries
```{r, echo=TRUE, results='hide', warning=FALSE, message=FALSE}
library(ggplot2)
library(lubridate)
library(dplyr)
```


## Loading the data
```{r LoadData}
if(!file.exists("activity.csv")){
        unzip("activity.zip")
}
data <- read.csv("activity.csv", stringsAsFactors = FALSE)   
```

## preprocessing the date format in data
```{r PreProcessingData}
data$date <- ymd(data$date)
str(data)
head(data)
```

## What is mean total number of steps taken per day?
1. Calculate the total number of steps taken per day. Ignore the NA Value for Steps
```{r}
steps <- data %>%
  filter(!is.na(steps)) %>%
  group_by(date) %>%
  summarize(steps = sum(steps)) %>%
  print
```
2. Plot histogram for total number of steps taken per day
```{r}
ggplot(steps, aes(x = steps)) +
  geom_histogram(fill = "firebrick", binwidth = 1000) +
  labs(title = "Histogram of Steps per day", x = "Steps per day", y = "Frequency")
```

3. Mean and median of the total number of steps taken per day
```{r}
meansteps <- mean(steps$steps)
mediansteps <- median(steps$steps)
meansteps
mediansteps
```
 

## What is the average daily activity pattern?
1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)?
```{r}
interval <- data %>%
  filter(!is.na(steps)) %>%
  group_by(interval) %>%
  summarize(steps = mean(steps))
ggplot(interval, aes(x=interval, y=steps)) +
  geom_line(color = "firebrick")
```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
interval[which.max(interval$steps),]
```


## Imputing missing values
1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
missingvalue <- data %>% filter(is.na(steps)) %>%
        count %>%
        print
```
2. Fill in all of the missing values in the dataset using mean of same 5 minute interval. 
```{r}
data_full <- data
nas <- is.na(data_full$steps)
```
3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
avg_interval <- tapply(data_full$steps, data_full$interval, mean, na.rm=TRUE, simplify=TRUE)
data_full$steps[nas] <- avg_interval[as.character(data_full$interval[nas])]
sum(is.na(data_full$steps))
```
4. Make a histogram of the total number of steps taken each day 
```{r}
steps_full <- data_full %>%
  filter(!is.na(steps)) %>%
  group_by(date) %>%
  summarize(steps = sum(steps)) %>%
  print
ggplot(steps_full, aes(x = steps)) +
  geom_histogram(fill = "firebrick", binwidth = 1000) +
  labs(title = "Histogram of Steps per day, including missing values", x = "Steps per day", y = "Frequency")
```

... and Calculate and report the mean and median total number of steps taken per day. 
```{r}
mean_steps_full <- mean(steps_full$steps, na.rm = TRUE)
median_steps_full <- median(steps_full$steps, na.rm = TRUE)
mean_steps_full
median_steps_full
```

The impact of imputing missing data ith the average number of steps in the same 5-min interval is that both the mean and the median are equal to the same value: `r mean_steps_full`



## Are there differences in activity patterns between weekdays and weekends?
For this part the weekdays() function may be of some help here. Use the dataset with the filled-in missing values for this part.

1. Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
```{r}
data_full <- mutate(data_full, weektype = ifelse(weekdays(data_full$date) == "Saturday" | weekdays(data_full$date) == "Sunday", "weekend", "weekday"))
data_full$weektype <- as.factor(data_full$weektype)
head(data_full)
```
2. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.
```{r}
data_full <- mutate(data_full, weektype = ifelse(weekdays(data_full$date) == "Saturday" | weekdays(data_full$date) == "Sunday", "weekend", "weekday"))
data_full$weektype <- as.factor(data_full$weektype)
head(data_full)
interval_full <- data_full %>%
        group_by(interval, weektype) %>%
        summarize(steps = mean(steps))

d <- ggplot(interval_full, aes(x=interval, y=steps, color=weektype))+
        geom_line() +
        facet_wrap(~weektype, ncol = 1, nrow = 2)
print(d)

```

